---
title: "PA1_template"
author: "Roberto Di Pietro"
date: "17/5/2019"
output: md_document
fig_caption: yes
keep_md: yes
toc: yes
pdf_document: default
keep_md: yes
self_contained: no
---

I first clean the Global Environment and set up the markdown page
```{r setup 1, include=FALSE}
rm(list=ls())
yourdirectory<-"F:/CasaUfficio/R notes/Reproduceable Research" #here you 
                                                                #plug in yours
require(knitr)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = yourdirectory)

```

I create a directory where I can download the data

```{r setup 2}
print("Hello, this is my version")
print(version)

#create dir
if(!file.exists("data")) {
        dir.create("data")
}

```

I download the data from the link
```{r download}

url<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
if(!file.exists("data.zip")) {
        download.file(url,destfile="data.zip")
        dateDownloaded<-date()
        print(dateDownloaded)
}

```

I unzip and open the data
```{r open}
mydir<-getwd()
print(mydir)
if(!file.exists("data.zip")) {
        unzip("data.zip", exdir=paste(mydir,"/data",sep=""))  
}
activities<-read.csv(paste(mydir,"/data/activity.csv",sep=""))
activities$date<-as.Date(as.character(activities$date))
```

I load the needed packages

```{r libraries}
library(dplyr)
library(ggplot2)
```


Request: Histogram of the total number of steps taken each day
Observations: the distribution is positive asymmetric, the bulk of the steps 
taken per day is between 10K and 15K, which is around 7 and 10 kilometers per 
day. I presume this user isn't exactly always home.. 

```{r 2}
activities %>% group_by(date) %>% 
        summarise("total steps"=sum(steps,na.rm = TRUE)) %>% 
        select("total steps") %>% pull() %>% hist()
```

Request:Mean and median number of steps taken each day

Observations: median is always back of the mean, confirming the skewnewss of the distribution of the steps. Median steps are often zero, this reads that in those
days the user is still (sleeping, laying or sat) for most of the day (+50% of 
the time).

```{r 3}
activities %>% group_by(date) %>% 
        summarise("mean steps"=mean(steps,na.rm = TRUE),
                  "median steps"=median(steps,na.rm = TRUE))
```

Request:Time series plot of the average number of steps taken

Observations: means per day reflect the different totals in number of steps 
taken per day: some days like in late november the user was very active, while 
on some other day the user seems to have not taken a single step out of bed (or 
most likely he forgot his smartwatch at home)

```{r 4}
avgstepsbyday<-activities %>% group_by(date) %>% 
        summarise("mean steps"=mean(steps,na.rm = TRUE)) %>% 
        filter(!is.nan(`mean steps`))
ggplot(avgstepsbyday,aes(date,`mean steps`))+geom_line()
```

Request: The 5-minute interval that, on average, contains the maximum number of 
steps

Observations: the interval in which the user takes more steps is no. 835.When 
dividing it by 60 one gets 13.9, which-supposing the intervals start their 
numbering at midnight- corresponds to a peak in activity some time before 2pm, 
that might be in his lunchbreak. 

```{r 5}
avgstepsbyinterval<-activities %>% group_by(interval) %>% 
        summarise(mean_steps=mean(steps,na.rm = TRUE)) 
maxstepinterval<-avgstepsbyinterval %>% filter(mean_steps==max(mean_steps)) 
maxstepinterval
```

Request: Code to describe and show a strategy for imputing missing data

Explanation: I chose to impute NAs to the median number of steps taken for the corresponding interval. 

```{r 6}
activitiesNAi<-activities 
for (i in unique(activities$interval)) {
  activitiesNAi$steps[is.na(activitiesNAi$steps) & activities$interval==i]<-
    activities %>% filter(interval==i) %>% select(steps) %>% pull() %>%
    median(steps,na.rm = T)   
  }

```

Request:Histogram of the total number of steps taken each day after missing 
values are 
imputed

Observation: given the choice of imputing missing to the medians, is all but 
strange that the prominence of the central bin was confirmed and reinforced

```{r 7}
activitiesNAi %>% group_by(date) %>% 
        summarise("total steps"=sum(steps,na.rm = TRUE)) %>% 
        select("total steps") %>% pull() %>% hist()
```

Request:Panel plot comparing the average number of steps taken per 5-minute 
interval 
across weekdays and weekends

Observation: I created two different panel, one that highlights the dynamics 
through the day of respectively the weekday and the weekend, and another that highlights the distribution of "step-intensity" of the intervals again in both 
weekdays and weeends.
What emerges is a higher mobility in weekends and a "more rigid" pattern of 
peaks of activity and still-intervals during weekdays. This is likely associated 
with a job that involves a mostly constant "routine" during weekdays than in 
weekends.

```{r 8}
activitiesWE<-activities %>% mutate(weekdays=weekdays(date))
activitiesWE$WE<-NA
activitiesWE$WE[activitiesWE$weekdays=="sabato" |
                  activitiesWE$weekdays=="domenica"] <-1
activitiesWE$WE[activitiesWE$weekdays!="sabato" &
                  activitiesWE$weekdays!="domenica"] <-0
avgstepsbyintervalbyWE<-activitiesWE %>% group_by(interval,WE) %>% 
        summarise(mean_steps=mean(steps,na.rm = TRUE))
avgstepsbyintervalbyWE$WE<-factor(avgstepsbyintervalbyWE$WE,levels=c(0,1),
                                  labels=c("weekday","weekend"))
ggplot(avgstepsbyintervalbyWE,aes(interval,mean_steps))+geom_line()+
      facet_grid(WE~.)
```


```{r 8b}
qplot(mean_steps,data=avgstepsbyintervalbyWE, fill=WE,binwidth=5 )
```
